NGO Analysis: `r params$title`

### 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
Testing automatic report generation.

If desired, a copy of the analysis in html format may then be downloaded.

```{r sidebar_button}
# download
uiOutput("downloadUI")
```

@@

output$downloadUI <- renderUI({
  downloadButton("report", "Generate report")
})

output$report <- downloadHandler(
  filename = "report.html",
  content = function(file) {
    tempReport <- file.path(tempdir(), "report.Rmd")
    file.copy("report.Rmd", tempReport, overwrite = TRUE)

    params <- list(
      htmlfile = input$html$datapath,
      excelfile = input$excel$datapath,
      start = input$start,
      end = input$end
    )

>>>>

If desired, a copy of the analysis in PDF format may then be downloaded.

Enter the name of the NGO in the text box to include it in the report.

```{r sidebar_button}
textInput("name", "NGO name", value = "NGO name")
# download
uiOutput("downloadUI")
```

@@

output$downloadUI <- renderUI({
  downloadButton("report", "Download report")
})

output$report <- downloadHandler(
  filename = "report.pdf",
  content = function(file) {
    tempReport <- file.path(tempdir(), "report.Rmd")
    file.copy("report.Rmd", tempReport, overwrite = TRUE)

    params <- list(
      htmlfile = input$html$datapath,
      excelfile = input$excel$datapath,
      start = input$start,
      end = input$end,
      title = input$name
    )

### 2022-07-24 Shiny/Test/report.Rmd
Testing automatic report generation.

---
title: "Report"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-24'
output: html_document
params:
  htmlfile: NA
  excelfile: NA
  start: NA
  end: NA
---

@@

if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using the fontawesome package for more up-to-date (and currently supported)
# Fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

# Functions
```{r}
# Data reading functions =======================================================

@@

exceldata <- widen_excel(params$excelfile)

# Support functions ============================================================

## scale_fill_green ------------------------------------------------------------
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}

@@

## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}

## Fundraising ROI------------------------------------------------------------
  # Creating Fundraising ROI value
    # testing for data quality, returning NA if bad
fundraising <- if (datatest == TRUE) {
      NA
    } else {
      exceldata |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    } |> KPI_to_string()

```

```

```{r data, include=FALSE}

```

@@

>>>>

---
title: "NGO Analysis: `r params$title`"
date: "`r Sys.Date()`"
output: pdf_document
params:
  htmlfile: NA
  excelfile: NA
  start: NA
  end: NA
  title: NA
---

@@

if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Data reading functions =======================================================

@@

exceldata <- widen_excel(params$excelfile)
```

```{r static_functions, echo=FALSE, message=FALSE, warning=FALSE}
## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}

@@

## scale_fill_green ------------------------------------------------------------
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}
```

```{r functions_reliant, echo=FALSE, message=FALSE, warning=FALSE}
## scale maximum--------------------------------------------------------------
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- 
    # reading the Excel sheet anew
    read_excel(params$excelfile) |>
      # removing the firm and sub-category columns
      select(-Firm, -`Sub-category`) |>
      # filtering to retain only revenue and expenses entries
      filter(Category == "Revenue" | Category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(Year, Category) |>
      summarise(
        sum = sum(Value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()

## Currency scale-------------------------------------------------------------
# Reactive function to check for currency and use currency code or ?
currency_scale <-
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )

## Plotly caption-------------------------------------------------------------
# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
# plotly_caption <-
#   list(
#     x = 1, y = -0.1, text = ifelse(length(unique(htmldata$valutakode)) == 1,
#       "", "Note: not all values are denominated in the same currency"
#     ),
#     showarrow = F, xref = "paper", yref = "paper",
#     xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
#     font = list(size = 11)
#   )

currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    )
```

```{r KPIs, echo=FALSE, message=FALSE, warning=FALSE}
## Fundraising ROI------------------------------------------------------------
# Creating Fundraising ROI value
# testing for data quality, returning NA if bad
fundraising <- if (datatest == TRUE) {
  NA
} else {
  {
    exceldata |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  } |> KPI_to_string()
}

## OPX KPI--------------------------------------------------------------------
# Creating OPX KPI
KPI_OPX_num <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX =
          sum(revenue_private_contributions,
            revenue_investments, assets_cash_in_bank_and_cash_equivalents,
            na.rm = TRUE
          ) /
            sum(liability_short_term_grants_payable,
              liability_revocable_endowments,
              liability_long_term_grants_payable,
              na.rm = TRUE
            ) * 100
      ) |>
      # Just the single number
      pull() |>
      first() |>
      KPI_to_string()
  }

# Creating private donations growth
KPI_private_num <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_private_contributions[1] -
            revenue_private_contributions[2]) /
            revenue_private_contributions[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first() |> 
      KPI_to_string()
  }

## Public grants growth-------------------------------------------------------
# Creating public grants growth
KPI_public_num <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_public_grants[1] - revenue_public_grants[2]) /
            revenue_public_grants[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first() |> 
      KPI_to_string()
  }

## Public/private proportion--------------------------------------------------
# Creating public/private proportion
KPI_proportion <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      transmute(
        proportion =
          sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) /
            sum(revenue_public_grants, revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) * 100
      ) |>
      pull() |>
      first() |>
      KPI_to_string()
  }
```

@@

```{r plots, echo=FALSE, message=FALSE, warning=FALSE}
# Plots ======================================================================

## Annual result--------------------------------------------------------------
arsresultat_plot <-
  # testing for the presence of data
  if (sum(!is.na(htmldata$arsresultat)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    htmldata |>
      # renaming for plotly hover
      rename(
        Year = year,
        `Net income` = arsresultat,
        `Net positive` = annual_profit
      ) |>
      ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$annual_profit) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Net income (in thousands)",
        caption = currency_caption
      )
  }

arsresultat_plot

## Balance--------------------------------------------------------------------
balance_plot <- 
  # testing for the presence of data
  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    htmldata |>
      # rename for plotly hover
      rename(
        Year = year,
        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
        `Positive equity` = balance_positive
      ) |>
      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$balance_positive) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Total equity/liabilities (in thousands)",
        caption = currency_caption
      )
  }

balance_plot

# Creating plotly output
#ggplotly(balance_plot) |>
  ## ggplot subtitles and captions are lost when using ggplotly(), a
  ## currency warning caption is therefore added in plotly, if needed, using
  ## a separate function
  # layout(
  #   annotations =
  #     plotly_caption
  # )

## Expenses-------------------------------------------------------------------
expenses_plot <- 
    # testing for data quality
    if (datatest == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum)
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }

expenses_plot

# Creating plotly output
# ggplotly(expenses_plot)

## Revenue--------------------------------------------------------------------
revenue_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    # printing 'bad data' plot if data are bad
    bad_data()
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)")
  }

revenue_plot

# Creating plotly output
# ggplotly(revenue_plot)

## CEO compensation-----------------------------------------------------------
CEO_plot <-
  # Testing for the presence of data
  if (sum(!is.na(htmldata$lederlonn)) == 0 &
    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    # Creating plot if any data exist
    # Note that the plot will still look empty, but have a legend and scales,
    # if all values are explicit zeroes
    htmldata |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # pretty colours
      scale_fill_met_d("Hokusai2", direction = -1) +
      labs(
        x = "", y = "Total compensation", title = "CEO compensation"
      )
  }

CEO_plot

# Creating plotly output
# ggplotly(CEO_plot) |>
  ## ggplot subtitles and captions are lost when using ggplotly(), a
  ## currency warning caption is therefore added in plotly, if needed, using
  ## a separate function
  # layout(
  #   annotations =
  #     plotly_caption
  # )
```
